\section{Lie algebras}

\begin{definition}
    \index{Lie algebra}
    Let $K$ be a field. 
    A (complex) \emph{Lie algebra} is a complex vector space
    $L$ together with a bilinear map 
    $L\times L\to L$, $(x,y)\mapsto [x,y]$,
    such that
    \begin{align}
        \label{eq:[xx]=0}&[x,x]=0\quad\text{for all $x\in L$},\\ 
        \label{eq:Jacobi}&[x,[y,z]]+[y,[z,x]]+[z,[x,y]]=0\quad\text{for all $x,y,z\in L$}.
    \end{align}
\end{definition}

\index{Jacobi identity}
Equality \eqref{eq:Jacobi} is known as the \emph{Jacobi identity}. 

\begin{exercise}
    Prove that \eqref{eq:[xx]=0} implies $[x,y]=-[y,x]$ for all
    $x,y\in L$. 
\end{exercise}

\index{Abelian Lie algebra}
A Lie algebra $L$ is said to be \emph{abelian} if $[x,y]=0$ for 
all $x,y\in L$. 

\begin{exercise}
    If $L$ and $L_1$ are Lie algebras, then 
    $L\oplus L_1$ is a Lie algebra with
    $[(x,x_1),(y,y_1)]=([x,y),(x_1,y_1)]$ for $x,y\in L$ and
    $x_1,y_1\in L_1$. 
\end{exercise}

\begin{exercise}
    Prove that $\R^3$ with the usual vector product 
    \[
    [(x_1,x_2,x_3),(y_1,y_2,y_3)]=(x_2y_3-x_3y_2,x_3y_1-x_1y_3,x_1y_2-x_2y_1)
    \]
    is a (real) Lie algebra.     
\end{exercise}

We will mainly work with finite-dimensional complex Lie algebras.

\begin{example}[general linear Lie algebra]
\index{General linear Lie algebra}
    Let $V$ be a finite-dimensional vector space and 
    $\gl(V)$ be the set of linear maps $V\to V$. Then 
    $\gl(V)$ with $[x,y]=xy-yx$ is a Lie algebra. 
\end{example}

A matrix version of the previous example: We write $\gl(n,\C)$ 
to denote the vector space of all $n\times n$ complex 
matrices with Lie bracket $[x,y]=xy-yx$. The vector space
$\gl(n,\C)$ has a basis $\{e_{ij}:1\leq i,j\leq n\}$, where
\[
(e_{ij})_{kl}=\begin{cases} 
    1 & \text{if $(i,j)=(k,l)$},\\
    0 & \text{otherwise}.
    \end{cases}
\]

\begin{exercise}
    Compute $[e_{ij},e_{ik}]$.
\end{exercise}

\begin{example}[special linear Lie algebra]
\index{Special linear Lie algebra}
    Let $\sl(n,\C)$ be the subspace of $\gl(n,\C)$ consisting
    of all matrices with trace zero. 
\end{example}

\begin{exercise}
    Find a basis of $\sl(n,\C)$. 
\end{exercise}

%\begin{example}
%    Let $\bl(n,\C)$ be the subspace of all upper triangular matrices
%    in $\gl(n,\C)$. Then $\bl(n,\C)$ is a Lie algebra. 
%\end{example}

\begin{definition}
    \index{Lie!subalgebra}
    A Lie \emph{subalgebra} of $L$ is a vector space $L_1$ of $L$ 
    such that $[x,y]\in L_1$ for all $x,y\in L_1$. 
\end{definition}

Of course, $\sl(n,\C)$ is a subalgebra of $\gl(n,\C)$. 

\begin{definition}
\index{Ideal!of a Lie algebra}
    An \emph{ideal} of a Lie algebra $L$ is a subspace $I$ of $L$ 
    such that $[x,y]\in I$ for all $x\in L$ and $y\in I$. 
\end{definition}

Trivial examples of ideals of a Lie algebra $L$ are
$\{0\}$ and $L$.

\begin{example}
\index{Center!of a Lie algebra}
    Let $L$ be a Lie algebra. Then 
    the \emph{center} 
    \[
    Z(L)=\{x\in L:[x,y]=0\text{ for all $y\in L$}\}.
    \]
    is an ideal of $L$. 
\end{example}

\begin{example}
\index{Derived algebra!of a Lie algebra}
    Let $L$ be a Lie algebra. 
    The \emph{derived algebra} $[L,L]$
    consists of all linear combinations of commutators $[x,y]$ 
    is an ideal of $L$. 
\end{example}

\begin{exercise}
    Compute $Z(\sl(n,\C))$. 
\end{exercise}

\begin{exercise}
    Prove that $\sl(2,\C)$ has no non-trivial ideals. 
\end{exercise}

One easily checks that $\sl(n,\C)$ is an ideal of $\gl(n,\C)$. In fact, 
an ideal is always a subalgebra. The converse is not true. 
Can you find an example?

\begin{definition}
\index{Homomorphism!of Lie algebras}
    Let $L$ and $L_1$ be Lie algebras. A map $f\colon L\to L_1$ is a 
    \emph{Lie algebra homomorphism} if $f([x,y])=[f(x),f(y)]$ for all
    $x,y\in L$. 
\end{definition}

As usual, an isomorphism between Lie algebras will be
a bijective homomorphism of Lie algebras. 

\begin{example}
    Let $L$ and $L_1$ be Lie algebras. The canonical injections
    $L\to L\oplus L_1$ and $L_1\to L_\oplus L_1$ and
    the canonical surjections $L\oplus L_1\to L$ and 
    $L\oplus L_1\to L_1$ are Lie algebras homomorphisms.  
\end{example}

\begin{example}
    Let $L$ be a Lie algebra. The \emph{opposite Lie algebra} 
    $L^{\op}$ is the vector space $L$ with 
    $[x,y]^{\op}=-[x,y]$. Then $L\to L^{\op}$, $x\mapsto -x$, 
    is an isomorphism of Lie algebras.
\end{example}

\begin{exercise}
    Let $f\colon L\to L_1$ be a Lie algebra homomorphism. Prove
    that the \emph{kernel} of $f$, 
    $\ker f=\{x\in L:f(x)=0\}$ is an ideal
    of $L$, and that the \emph{image} of $f$ 
    is a subalgebra of $L_1$. 
\end{exercise}

\begin{example}
\index{Adjoint homomorphism}
    Let $L$ be a Lie algebra. 
    The \emph{adjoint homomorphism} is the map 
    \[
    \ad\colon L\to\gl(L),\quad
    (\ad x)(y)=[x,y].
    \]
\end{example}

Let $L$ be a Lie algebra and $I$ be an ideal of $L$. Then 
the quotient vector space $L/I$ is a Lie algebra
with $[x+I,y+I]=[x,y]+I$. The canonical map 
$L\to L/I$, $x\mapsto x+I$, 
is a surjective Lie algebra homomorphism. 

\begin{exercise}
    Let $f\colon L\to L_1$ be a Lie algebra homomorphism.
    Prove that $f/\ker f\simeq f(L)$. 
\end{exercise}

\begin{definition}
\index{Simple Lie algebra}
    A Lie algebra $L$ is said to be \emph{simple} if 
    $[L,L]\ne\{0\}$ and $\{0\}$ and $L$ are the only ideals of $L$. 
\end{definition}

If $L$ is a simple Lie algebra, then $Z(L)=\{0\}$ and $L=[L,L]$. 

\begin{exercise}
    Prove that every simple Lie algebra is isomorphic to 
    a linear Lie algebra. 
\end{exercise}

\subsection{Representations of Lie algebras}

\begin{definition}
\index{Representation!of a Lie algebra}
    A \emph{representation} of a Lie algebra $L$ 
    is a Lie homomorphism $\rho\colon L\to\gl(V)$, where $V$ is a vector space. 
\end{definition}

If $L\to\gl(V)$ is a representation of a Lie algebra $L$, 
fixing a basis for $V$ 
we obtain a \emph{matrix representation}
$L\to\gl(n,\C)$. 

\begin{example}
    Let $L$ be a Lie algebra. 
    The map $\ad\colon L\to\gl(L)$, $x\mapsto (\ad x)$, is a lie 
    homomorphism. 
\end{example}

\begin{definition}
Let $L$ be a Lie algebra. 
A (left) Lie $L$-module is a vector space $V$ 
together with a map $L\times V\to V$, $(x,v)\mapsto xv$, 
such that $(x,v)\mapsto xv$ is bilinear 
and 
\[
    [x,y]v=x(yv)-y(xv)
\]
for all $x,y\in L$ and $v\in V$. 
\end{definition}

As it happens in the case of groups, Lie modules are
in bijective correspondence with representations. 

\begin{example}
    Let $L$ be a subalgebra of $\gl(V)$. Then 
    $L$ is an $L$-module. 
\end{example}

\begin{definition}
    Let $L$ be a Lie algebra and $V$ be a Lie $L$-module. 
    A \emph{submodule} of $V$ is a subspace $W$ 
    such that $xw\in W$ for all $x\in L$ and $w\in W$. 
\end{definition}

\begin{example}
We know that $L$ is an $L$-module 
with the adjoint representation. The submodules of $L$ are
the ideals of $L$. 
\end{example}

If $W$ is a submodule of $V$, then $V/W$ 
with $x(v+W)=xv+W$ is a module. 

\begin{definition}
    Let $L$ be a Lie algebra. An $L$-module $V$ 
    is said to be \emph{simple} (or irreducible) 
    if $V\ne \{0\}$ and it has no submodules other than $\{0\}$ and $V$. 
\end{definition}

One-dimensional modules are simple. In particular, 
the trivial module is always simple. 

\begin{example}
    Let $L$ be a simple Lie algebra (e.g. $\sl(2,\C)$). Then 
    the adjoint representation is irreducible, that is $L$ is a simple $L$-module.  
\end{example}

\begin{definition}
    Let $L$ be a Lie algebra and $V$ be an $L$-module. We say 
    that $V$ is \emph{indecomposable} if 
    there are no non-zero submodules $U$ and $W$ such that
    $V=U\oplus W$.
\end{definition}

Clearly, irreducible modules are indecomposable. 
The converse is not true.


\begin{exercise}
    Let $\mathfrak{b}(n,\C)$ be the set of $n\times n$ 
    upper triangular 
    matrices in $\gl(n,\C)$. Prove that $V=\C^n$ is indecomposable,  
    not irreducible. 
\end{exercise}

\begin{definition}
\index{Homomorphism!of modules over Lie algebras}
    Let $L$ be a Lie algebra and $f\colon V\to W$ be a map. 
    We say that $f$ is an $L$-module \emph{homomorphism} if
    $f(xv)=xf(v)$ for all $x\in L$ and $v\in V$. 
\end{definition}

As usual, an isomorphism is a bijective module homomorphism. 

\begin{exercise}
    State and prove the isomorphism theorems for modules
    over Lie algebras. 
\end{exercise}

\begin{exercise}[Schur's lemma]
\index{Schur's lemma for Lie algebras}
    Let $L$ be a Lie algebra.
    \begin{enumerate}
        \item Let $S$ and $T$ be simple $L$-modules.
            Prove that a non-zero homomorphism $f\colon S\to T$ 
            is an isomorphism.
        \item Let $S$ be a finite-dimensional simple $L$-module. 
            Prove that if $f\colon S\to S$ is a homomorphism, then
                $f=\lambda\id$ for some $\lambda\in\C$. 
    \end{enumerate} 
\end{exercise}

As an example, if $V$ is a simple module, then $z$
acts by scalar multiplication on $V$, that is
$zv=\lambda v$ for some $\lambda\in\C$. 
